\Large \bf Cyclone: Replication Middleware for NVM Applications

### 2016-09-20 paper/paper.tex
add packet layout figure to paper.
Additions to paper.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\begin{document}

@@

\subsection{Zero Copy Replication}
Consensus protcols such as RAFT usually have a simple common case path. Once a
quorum leader is elected, the leader's primary job is to transmit new entries
appended to its persistent log to its followers. The primary job of a follower
is to receive these entries, append them to its persistent log and send
acknowledgments to the leader. An entry is considered committed when the leader
receives an acknowledgment for an entry from a majority of the quorum. Followers
receive intimation of committed entries as piggybacked information on subsequent
replication messages. The common case path at the leader lends itself to zero
copy since the same set of log entries is sent to all replicas.

@@

We exploit the fact that DPDKs packet buffers allow prepending data to a packet
without physically moving the data in the packet buffer. Fig
[figures/packets.bmp] shows how the layout of a packet buffer changes at the
leader during replication. The received packets contains only the RPC call from
the client. We prepend a RAFT related header (containing the previous log index,
previous log term, leader's term and leader's commit point) to packet buffer
itself. To this RAFT header we then prepend an IPv4 header. We set IP addresses
in the header to use 5-tuple steering in order to place the packet in the right
queue at the receiving NIC. This composite packet buffer is then sent to every
replica. In order to properly route the packet through switches, we allocate a
small buffer containing an ethernet header for each follower replica and
\emph{chain} it in front of the prepared packet - ensuring we do not need to
change the contents of the packet itself for every replica.

\subsection{Batching}
Although the common case path in RAFT is straightforward, it still adds many
hundreds of cycles for each log entry to be processed. This adds up to a few
microseconds, a significant amount in the face of sub 10 us network latency
between nodes with DPDK. We amortize this overhead when operating under load by
batching multiple RPC calls together and treating the batch as a single RAFT log
entry.  DPDK by itself can efficiently batch incoming packets under load -
allowing the user to easily receive a burst (currently limited to 32) of packets
rather than a single one for a polling call. DPDK further allows chaining packet
buffers together and treating the chain as a single packet. This is convenient
for our zero copy approach and we chain packet mbufs together to form a single
packet as shown in Fig [figures/chain.bmp]. Since the entire chain is sent as a
single packet when being replicated, the length of a chain is limited by the
size of the coalesced packet, that must fit into an ethernet jumbo frame of
approximately 9KB.

>>>>

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,graphicx}
\begin{document}

@@

\subsection{Zero Copy Batching}
Consensus protcols such as RAFT usually have a simple common case path. Once a
quorum leader is elected, the leader's primary job is to transmit new entries
appended to its persistent log to its followers. The primary job of a follower
is to receive these entries, append them to its persistent log and send
acknowledgments to the leader. An entry is considered committed when the leader
receives an acknowledgment for an entry from a majority of the quorum. Followers
receive intimation of committed entries as piggybacked information on subsequent
replication messages. The common case path at the leader lends itself to zero
copy since the same set of log entries is sent to all replicas.

@@

We exploit the fact that DPDKs packet buffers allow prepending data to a packet
without physically moving the data in the packet buffer. Fig
[figures/packets.bmp] shows how the layout of a packet buffer changes at the
leader during replication. The received packets contains only the RPC call from
the client. We prepend a RAFT related header (containing the previous log index,
previous log term, leader's term and leader's commit point) to packet buffer
itself. To this RAFT header we then prepend an IPv4 header. We set IP addresses
in the header to use 5-tuple steering in order to place the packet in the right
queue at the receiving NIC. We do not use different ip addresses for different
replicas, rather the ip addresses are only set in conjunction with NIC
programming to enable steering of the received packets to the correct NIC queue.
This composite packet buffer is then sent \emph{unchanged} to every replica. In
order to properly route the packet through switches, we allocate a small buffer
containing an ethernet header for each follower replica and \emph{chain} it in
front of the prepared packet - ensuring we do not need to change the contents of
the packet itself for every replica.

Although the common case path in RAFT is straightforward, it still adds many
hundreds of cycles for each log entry to be processed. This adds up to a few
microseconds, a significant amount in the face of sub 10 us network latency
between nodes with DPDK. We amortize this overhead when operating under load by
batching multiple RPC calls together and treating the batch as a single RAFT log
entry.  DPDK by itself can efficiently batch incoming packets under load -
allowing the user to easily receive a burst (currently limited to 32) of packets
rather than a single one for a polling call. DPDK further allows chaining packet
buffers together and treating the chain as a single packet. This is convenient
for our zero copy approach and we chain packet mbufs together to form a single
packet as shown in Fig [figures/chain.bmp]. Since the entire chain is sent as a
single packet when being replicated, the length of a chain is limited by the
size of the coalesced packet, that must fit into an ethernet jumbo frame of
approximately 9KB.

\begin{figure}
\includegraphics[width=0.7\textwidth,height=6cm]{figures/chain.pdf}
\caption{Zero Copy Batching}
\label{fig:zc_batch}
\end{figure}

This set of transformations to incoming request packets from clients at the
leader replica is summarized in Figure~\ref{fig:zc_batch}.
